---
title: "glm Jawa Tengah - skripsi"
author: "cindy 01112180008"
date: "9/19/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

``````{r, echo=TRUE}
library(dplyr)
library(faraway)
library(MASS)
library(Metrics)
library(car)
library(AER)
```

# 1. data input for normality and dispersion test
``````{r, echo=TRUE}
data = data_covid_seluruh_provinsi
```

#normality test
``````{r, echo=TRUE}
shapiro.test(data$`jawa tengah`)
```

#dispersion test
``````{r, echo=TRUE}
dispersion_test(data$`jawa tengah`)
```

# 2. Counting GLM model

#input and split data
``````{r, echo=TRUE}
data = jawa_tengah
data=within(data,{
  rm(Tanggal)
})
# split data
set.seed(1)
data_size=0.8*nrow(data)
sample <- sample.int(n = nrow(data), data_size, replace = F)
jateng_train=data[sample,]
jateng_test=data[-sample,]
jateng_train
jateng_test
```

#create Poisson GLM
``````{r, echo=TRUE}
poisson_jateng_train = glm(jateng_train$Penambahan_Harian_Kasus_Terkonf~., family=poisson(link="log"), data = jateng_train)
summary(poisson_jateng_train)
# poisson glm for training data
```

#look for VIF values to see if there are any multicollinearity between variables
``````{r, echo=TRUE}
vif(poisson_jateng_train)
# all VIF < 10, no multicollinearity between variables. All variables used to create Poisson model
```

#check for overdispersion in poisson glm with pearson goodness of fit
``````{r,echo=TRUE}
x2 = sum(residuals(poisson_jateng_train, type="pearson")^2)
c(Df=df.residual(poisson_jateng_train), resid.dev=deviance(poisson_jateng_train),peason.x2 = x2)
pchisq(deviance(poisson_jateng_train),df.residual(poisson_jateng_train),lower.tail=FALSE)
# deviance(poisson_jateng_train)/df.residual(poisson_jateng_train)
# residual deviance & pearson goodness-of-fit > residual df -> data overdispersed (model inadequate)
# p-value = 0 -> reject goodness of fit test, model inadequate
```

#create negative binomial GLM
``````{r, echo=TRUE}
nb_jateng_train = glm.nb(jateng_train$Penambahan_Harian_Kasus_Terkonf~., data = jateng_train, control=glm.control(maxit=200))
summary(nb_jateng_train)
```

#look for VIF values to see if there are any multicollinearity between variables
``````{r, echo=TRUE}
vif(nb_jateng_train)
```

#create nb model with backward selection
``````{r, echo=TRUE}
nb_jateng_train_2= stepAIC(nb_jateng_train, direction = c("backward"), trace = FALSE)
summary(nb_jateng_train_2)
vif(nb_jateng_train_2)
```

``````{r, echo=TRUE}
# exp(14.21054)
# exp(-0.13885)
# exp(0.11968)
# exp(0.23343)
# exp(0.05900)
# exp(0.02842)
# exp(-0.03633)
```

#predict with negative binomial GLM
``````{r, echo=TRUE}
jateng_test$prediction = as.integer(predict(nb_jateng_train_2, newdata=jateng_test, type="response"))
jateng_test
```

#MAE values of the selected GLM
``````{r, echo=TRUE}
mae(jateng_test$Penambahan_Harian_Kasus_Terkonf,jateng_test$prediction)
mape(jateng_test$Penambahan_Harian_Kasus_Terkonf,jateng_test$prediction)
mase(jateng_test$Penambahan_Harian_Kasus_Terkonf,jateng_test$prediction)
rmse(jateng_test$Penambahan_Harian_Kasus_Terkonf,jateng_test$prediction)
```
